---
title: "ORCHIDS_NMDS_NO_SWALE"
author: "Joe Gunn"
date: "5/21/2020"
output: html_document
---

## Libraries needed for analysis
```{r libraries, echo=FALSE}
library(readxl)
library(cowplot)
library(tidyverse)
library(vegan)
library(nlme)
library(lme4)
library(factoextra)
library(viridis)
library(bbmle)
library(car)
library(multcomp)
library(RColorBrewer)
library(devtools)
library(pairwiseAdonis)
library(EcolUtils)
```

## Metadata
```{r}
#Read in site-level metadata and clean
site_metadata <- read_excel("../../data/metadata/site_metadata.xlsx")

site_metadata <- site_metadata %>%
  mutate(Site = factor(Site), Near_Far = factor(Near_Far), Location = factor(Location), loc_abbr = factor(loc_abbr)) %>%
  mutate(present_absent = ifelse(as.factor(Near_Far) == "NO" | as.factor(Near_Far) == "NOE", "Absent", "Present"))

#Read in taxonomic level metadata and clean
tax_metadata <- read_excel("../../data/metadata/tax_metadata.xlsx")

tax_metadata <- tax_metadata %>%
  mutate(kingdom = factor(kingdom),
         phylum = factor(phylum), 
         class = factor(class), 
         order = factor(order), 
         family = factor(family), 
         genus = factor(genus), 
         species = factor(species))
```

## Read-in raw data
```{r}
#Read in OTU raw abundance data. This data is organized as a matrix of individual microbial OTUS (rows) and sampling sites (columns). 
soil_raw <- read_excel("../../data/excel_data/raw_data.xlsx")

#Clean table
colnames(soil_raw)[1] <- "otu_id"

soil_raw <- soil_raw %>%
  mutate(otu_id = as.character(otu_id)) %>%
  mutate(otu_id = factor(otu_id))

#Summarize raw microbial OTU data
#Get number of OTUs
otu_count <- soil_raw %>%
  nrow() #84,960 OTUs

#Get number of sites
site_count <- soil_raw %>%
  ncol() #24 sites (excluding otu id)

read_count <- soil_raw[,-c(1)] %>%
  colSums() #54,631 reads per sample - THE DATA IS ALREADY RAREFIED!


#combine taxonomy data (tax_metadatat) and raw otu data (soil_raw)
soil_all <- cbind(tax_metadata, soil_raw)
```

## Get Relative Abundance Data at the Phylum Level in a gathered dataset - this was produced in ORCHIDS_tax_confidence.rmd
```{r}
soil_data_phylum_raw <- soil_all[,-c(1,3:7)]

soil_data_phylum_rel <- soil_data_phylum_raw %>%
  mutate(LT0116S = LT0116S/54631,
         LT0216S = LT0216S/54631,
         NE0616S = NE0616S/54631,
         P0116S = P0116S/54631,
         P0216S = P0216S/54631,
         P0316S = P0316S/54631,
         H0316S = H0316S/54631,
         H0216S = H0216S/54631,
         H0116S = H0116S/54631,
         LT0516S = LT0516S/54631,
         T0416S = T0416S/54631,
         NE0216S = NE0216S/54631,
         NE0416S = NE0416S/54631,
         S0216S = S0216S/54631,
         LT0416S = LT0416S/54631,
         NE0516S = NE0516S/54631,
         S0116S = S0116S/54631,
         LT0316S = LT0316S/54631,
         NE0316S = NE0316S/54631,
         NE0116S = NE0116S/54631,
         T0316S = T0316S/54631,
         NE0816S = NE0816S/54631,
         NE0716S = NE0716S/54631,
         T0116S = T0116S/54631)

#Full Dataset with raw and relative abundance data
soil_data_phylum_top10_raw_abundance_general <- soil_data_phylum_raw %>% 
  filter(phylum == "Verrucomicrobia" | phylum == "Proteobacteria" | phylum == "Actinobacteria" | phylum == "Acidobacteria" | phylum == "Chloroflexi" | phylum == "Planctomycetes" | phylum == "Bacteroidetes" | phylum == "Firmicutes" | phylum == "Nitrospirae" | phylum == "Gemmatimonadetes")
  
soil_data_phylum_top10_relative_abundance_general <- soil_data_phylum_rel %>% 
  filter(phylum == "Verrucomicrobia" | phylum == "Proteobacteria" | phylum == "Actinobacteria" | phylum == "Acidobacteria" | phylum == "Chloroflexi" | phylum == "Planctomycetes" | phylum == "Bacteroidetes" | phylum == "Firmicutes" | phylum == "Nitrospirae" | phylum == "Gemmatimonadetes")
```

## Get raw abundance and relative abundance datasets that are ready to use for all analyses
```{r}
soil_data_phylum_top10_raw_abundance <- column_to_rownames(soil_data_phylum_top10_raw_abundance_general, "otu_id")
soil_data_phylum_top10_relative_abundance <- column_to_rownames(soil_data_phylum_top10_relative_abundance_general, "otu_id")
```

## Get separate datasets without SWALE
```{r}
#Datasets with "Far" samples removed
soil_data_phylum_top10_raw_abundance_noSWALE <- soil_data_phylum_top10_raw_abundance[,-c(15,18)]
soil_data_phylum_top10_relative_abundance_noSWALE <- soil_data_phylum_top10_relative_abundance[,-c(15,18)]
```

## Run NMDS on all OTUs together and assess stress
```{r}
#Clean the dataset for all OTUs at the phylum leel to prepare data for NMDS
all_raw_abundance_distance_noSWALE <- soil_data_phylum_top10_raw_abundance_noSWALE[,-c(1)]
all_raw_abundance_distance_noSWALE <- as.matrix(t(all_raw_abundance_distance_noSWALE))

#Tranform data by total abundance across all sites
all_transform_noSWALE <- decostand(all_raw_abundance_distance_noSWALE, "total")

#Get distance matrix using Bray-Curtis distance
all_dist_noSWALE <- vegdist(all_transform_noSWALE, "bray")

#Run NMDS for all OTUs for 1, 2, and 3 dimensions
all_nmds_noSWALE_1 <- metaMDS(all_dist_noSWALE, k = 1, autotransform = F, expand = F, trace = F)
all_nmds_noSWALE_2 <- metaMDS(all_dist_noSWALE, k = 2, autotransform = F, expand = F, trace = F)
all_nmds_noSWALE_3 <- metaMDS(all_dist_noSWALE, k = 3, autotransform = F, expand = F, trace = F)

#Assess stress for each NMDS level
stress_noSWALE_1 <- all_nmds_noSWALE_1$stress
stress_noSWALE_2 <- all_nmds_noSWALE_2$stress
stress_noSWALE_3 <- all_nmds_noSWALE_3$stress

all_k_stress_noSWALE <- data.frame(cbind("k" = c(seq(1:3)), "stress" = c(stress_noSWALE_1, stress_noSWALE_2, stress_noSWALE_3)))

pdf("/Users/joegunn/Desktop/Grad_School_Stuff/Research/Projects/Soil_Microbiome/visualization/all_nmds_figures/no_swale/all_otus_nmds_stress_plot_noSWALE.pdf", width = 8, height = 5)

ggplot(all_k_stress_noSWALE, aes(x = k, y = stress)) +
  geom_point(size = 3) +
  geom_line() +
  theme_set(theme_cowplot(12)) +
  labs(x = "Number of NMDS Dimensions", y = "Stress") + 
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 15))

dev.off()
```

## Build NMDS plots
```{r}
#Extract NMDS scores and clean the dataset to plot
all_scores_noSWALE <- as.data.frame(scores(all_nmds_noSWALE_2))

all_scores_noSWALE <- all_scores_noSWALE %>% 
  rownames_to_column("Site") %>% 
  mutate(Site = factor(Site)) %>%
  left_join(site_metadata, "Site")

#Get centroid values and segment values from score data to make spider plot
gg_orchids_location_noSWALE <- merge(all_scores_noSWALE, aggregate(cbind(mean.x = NMDS1, mean.y = NMDS2) ~ Location, all_scores_noSWALE, mean), by = "Location")
gg_orchids_distance_noSWALE <- merge(all_scores_noSWALE, aggregate(cbind(mean.x = NMDS1, mean.y = NMDS2) ~ Near_Far, all_scores_noSWALE, mean), by = "Near_Far")
gg_orchids_present_absent_noSWALE <- merge(all_scores_noSWALE, aggregate(cbind(mean.x = NMDS1, mean.y = NMDS2) ~ present_absent, all_scores_noSWALE, mean), by = "present_absent")


#Make NMDS plot for Location
pdf("/Users/joegunn/Desktop/Grad_School_Stuff/Research/Projects/Soil_Microbiome/visualization/all_nmds_figures/no_swale/all_otus_nmds_location_noSWALE.pdf", width = 8, height = 5)

ggplot(gg_orchids_location_noSWALE, aes(x = NMDS1, y = NMDS2, color = loc_abbr)) +
  geom_point(aes(color = loc_abbr), size = 3, show.legend = T) +
  geom_point(aes(x = mean.x, y = mean.y), color = "black", pch = 21, size = 3, show.legend = T) + 
  geom_segment(aes(x = mean.x, y = mean.y, xend = NMDS1, yend = NMDS2), show.legend = F) +
  labs(x = "NMDS 1", y = "NMDS 2", color = "Location") + 
  theme_set(theme_cowplot(12)) +
  scale_color_manual(values = c("#1B9E77","#D95F02","#7570B3","#E7298A","#E6AB02"))+
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(legend.position = c(0.05,0.2)) +
  theme(legend.title = element_text(size = 15)) +
  theme(legend.title = element_text(face = "bold")) + 
  theme(legend.text = element_text(size = 15)) +
  theme(legend.background = element_rect(color = "black")) +
  theme(legend.margin = margin(3,3,3,3))

dev.off()

#Make NMDS plot for Distance from Orchid
pdf("/Users/joegunn/Desktop/Grad_School_Stuff/Research/Projects/Soil_Microbiome/visualization/all_nmds_figures/no_swale/all_otus_nmds_distance_noSWALE.pdf", width = 8, height = 5)

ggplot(gg_orchids_distance_noSWALE, aes(x = NMDS1, y = NMDS2, color = Near_Far)) +
  geom_point(aes(color = Near_Far), size = 3, show.legend = T) +
  geom_point(aes(x = mean.x, y = mean.y), color = "black", pch = 21, size = 3, show.legend = T) + 
  geom_segment(aes(x = mean.x, y = mean.y, xend = NMDS1, yend = NMDS2), show.legend = F) +
  labs(x = "NMDS 1", y = "NMDS 2", color = "Distance from Orchid") + 
  theme_set(theme_cowplot(12)) +
  scale_color_brewer(palette = "Set1") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(legend.position = c(0.05,0.2)) +
  theme(legend.title = element_text(size = 15)) +
  theme(legend.title = element_text(face = "bold")) + 
  theme(legend.text = element_text(size = 15)) +
  theme(legend.background = element_rect(color = "black")) +
  theme(legend.margin = margin(3,3,3,3))

dev.off()

#Make NMDS plot for Orchids present or Absent
pdf("/Users/joegunn/Desktop/Grad_School_Stuff/Research/Projects/Soil_Microbiome/visualization/all_nmds_figures/no_swale/all_otus_nmds_present_absent_noSWALE.pdf", width = 8, height = 5)

ggplot(gg_orchids_present_absent_noSWALE, aes(x = NMDS1, y = NMDS2, color = present_absent)) +
  geom_point(aes(color = present_absent), size = 3, show.legend = T) +
  geom_point(aes(x = mean.x, y = mean.y), color = "black", pch = 21, size = 3, show.legend = T) + 
  geom_segment(aes(x = mean.x, y = mean.y, xend = NMDS1, yend = NMDS2), show.legend = F) +
  labs(x = "NMDS 1", y = "NMDS 2", color = "Orchids") + 
  theme_set(theme_cowplot(12)) +
  scale_color_manual(values = c("navyblue", "darkorchid1")) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(legend.position = c(0.05,0.2)) +
  theme(legend.title = element_text(size = 15)) +
  theme(legend.title = element_text(face = "bold")) + 
  theme(legend.text = element_text(size = 15)) +
  theme(legend.background = element_rect(color = "black")) +
  theme(legend.margin = margin(3,3,3,3))

dev.off()

##Make plots with object names so that you can put them into a plot_grid
nmds_location_noswale <- ggplot(gg_orchids_location_noSWALE, aes(x = NMDS1, y = NMDS2, fill = loc_abbr)) +
  geom_point(aes(x = NMDS1, y = NMDS2, fill = loc_abbr), pch = 21, color = "black", size = 3, show.legend = T) +
  geom_point(aes(x = mean.x, y = mean.y, fill = loc_abbr), pch = 21, color = "black", size = 3, show.legend = T) + 
  geom_segment(aes(x = mean.x, y = mean.y, xend = NMDS1, yend = NMDS2, color = loc_abbr), show.legend = F) +
  labs(x = "NMDS 1", y = "NMDS 2", color = "Location", fill = "Location") + 
  theme_set(theme_cowplot(12)) +
  scale_color_manual(values = c("#1B9E77","#D95F02","#7570B3","#E7298A","#E6AB02")) +
  scale_fill_manual(values = c("#1B9E77","#D95F02","#7570B3","#E7298A","#E6AB02")) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(legend.position = c(0.05,0.2)) +
  theme(legend.title = element_text(size = 15)) +
  theme(legend.title = element_text(face = "bold")) + 
  theme(legend.text = element_text(size = 15)) +
  theme(legend.background = element_rect(color = "black")) +
  theme(legend.margin = margin(3,3,3,3))

nmds_distance_noswale <- ggplot(gg_orchids_distance_noSWALE, aes(x = NMDS1, y = NMDS2, color = Near_Far)) +
  geom_point(aes(x = NMDS1, y = NMDS2, fill = Near_Far), pch = 21, color = "black", size = 3, show.legend = T) +
  geom_point(aes(x = mean.x, y = mean.y, fill = Near_Far), pch = 21, color = "black", size = 3, show.legend = T) + 
  geom_segment(aes(x = mean.x, y = mean.y, xend = NMDS1, yend = NMDS2, color = Near_Far), show.legend = F) +
  labs(x = "NMDS 1", y = "NMDS 2", color = "Distance from Orchid", fill = "Distance from Orchid") + 
  theme_set(theme_cowplot(12)) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(legend.position = c(0.05,0.2)) +
  theme(legend.title = element_text(size = 15)) +
  theme(legend.title = element_text(face = "bold")) + 
  theme(legend.text = element_text(size = 15)) +
  theme(legend.background = element_rect(color = "black")) +
  theme(legend.margin = margin(3,3,3,3)) +
  theme(axis.title.y = element_blank()) +
  theme(axis.text.y = element_blank())

nmds_pa_noswale <- ggplot(gg_orchids_present_absent_noSWALE, aes(x = NMDS1, y = NMDS2, fill = present_absent)) +
  geom_point(aes(x = NMDS1, y = NMDS2, fill = present_absent), pch = 21, color = "black", size = 3, show.legend = T) + 
  geom_point(aes(x = mean.x, y = mean.y, fill = present_absent),  pch = 21, color = "black", size = 3, show.legend = T) + 
  geom_segment(aes(x = mean.x, y = mean.y, xend = NMDS1, yend = NMDS2, color = present_absent), show.legend = F) +
  labs(x = "NMDS 1", y = "NMDS 2", color = "Orchid History", fill = "Orchid History") + 
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("navyblue", "darkorchid1")) +
  scale_color_manual(values = c("navyblue", "darkorchid1")) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(legend.position = c(0.05,0.14)) +
  theme(legend.title = element_text(size = 15)) +
  theme(legend.title = element_text(face = "bold")) + 
  theme(legend.text = element_text(size = 15)) +
  theme(legend.background = element_rect(color = "black")) +
  theme(legend.margin = margin(3,3,3,3)) +
  theme(axis.title.y = element_blank()) +
  theme(axis.text.y = element_blank())

pdf("/Users/joegunn/Desktop/Grad_School_Stuff/Research/Projects/Soil_Microbiome/visualization/all_nmds_figures/no_swale/all_otus_nmds_plots_noswale.pdf", width = 12, height = 5)

plot_grid(nmds_location_noswale, nmds_pa_noswale, nrow = 1, labels = c("a","b"), label_size = 25, label_y = 1, label_x = 0.02)

dev.off()
```

## Run PERMANOVA for all locations
## Get metadata in a format where it can be used with the orchid distance matrix
```{r}
##Here I am putting the metadata in the SAME ORDER as the distance matrix prodcued by vegan, because the order of the labels actually does appear to matter!
soil_meta_gathered_for_nmds_noSWALE <- soil_data_phylum_top10_raw_abundance_noSWALE %>%
  gather(LT0116S:T0116S, key = "Site", value = "raw_abundance")

soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[,c(2,3)]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(1:79587),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(2:79588),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(3:79589),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(4:79590),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(5:79591),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(6:79592),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(7:79593),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(8:79594),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(9:79595),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(10:79596),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(11:79597),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(12:79598),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(13:79599),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(14:79600),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(15:79601),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(16:79602),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(17:79603),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(18:79604),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(19:79605),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(20:79606),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(21:79607),]
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE[-c(22:79608),]

soil_meta_gathered_for_nmds_noSWALE <- as.data.frame(soil_meta_gathered_for_nmds_noSWALE)
soil_meta_gathered_for_nmds_noSWALE <- soil_meta_gathered_for_nmds_noSWALE %>% rownames_to_column("row")
soil_meta_gathered_for_nmds_noSWALE <- as.data.frame(soil_meta_gathered_for_nmds_noSWALE[,-c(1,3)])
colnames(soil_meta_gathered_for_nmds_noSWALE) <- "Site"

soil_metadata_permanova_noSWALE <- soil_meta_gathered_for_nmds_noSWALE %>%
  left_join(site_metadata, by = "Site")
```

## Run PERMANOVA
```{r}
set.seed(1234)

#Run Permanova for locations
permanova_all_locations_noSWALE <- adonis(all_dist_noSWALE ~ Location, soil_metadata_permanova_noSWALE, distance = "bray", permutations = 9999) # VERY SIGNIFICANT (p < 0.001)

hist(permanova_all_locations$f.perms)

#Run Permanova for distances from orchids
permanova_all_distance_noSWALE <- adonis(all_dist_noSWALE ~ Near_Far, soil_metadata_permanova_noSWALE, Strata = Location, distance = "bray", permutations = 9999) # NOT SIGNIFICANT 0.2019

hist(permanova_all_distance$f.perms)

#Run Permanova for present/absent orchids
permanova_orchids_present_absent_noSWALE <- adonis(all_dist_noSWALE ~ present_absent, soil_metadata_permanova_noSWALE, Strata = Near_Far, distance = "bray", permutations = 9999) # really close to significant 0.059, F = 1.741

hist(permanova_orchids_present_absent$f.perms)
```

## Get Beta Dispersion
```{r}
beta_dispersion_location <- betadisper(all_dist_noSWALE, soil_metadata_permanova_noSWALE$Location, type = "centroid")
anova(beta_dispersion_location) #Significant p = 0.034, F = 3.347

beta_dispersion_pa <- betadisper(all_dist_noSWALE, soil_metadata_permanova_noSWALE$present_absent, type = "centroid")
anova(beta_dispersion_pa) #Not significant p = 0.7752, F = 0.084
```