---
title: "ORCHIDS_UNUSED"
author: "Joe Gunn"
date: "4/17/2020"
output: html_document
---

## Test for significant differences at the Phylum Level between locations and between distances for each of the 10 most abundant phyla using NMDS 
### all analyses were conducted in the package vegan using NMDS methods
```{r}
###All datasets below are for individual phyla (top 10) and show raw abundance data
acidobacteria_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Acidobacteria") 

acidobacteria_raw_abundance_distance <- acidobacteria_raw_abundance_distance[,-c(1)]
acidobacteria_raw_abundance_distance <- column_to_rownames(acidobacteria_raw_abundance_distance, "otu_id")
acidobacteria_raw_abundance_distance <- as.matrix(t(acidobacteria_raw_abundance_distance))

actinobacteria_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Actinobacteria")

actinobacteria_raw_abundance_distance <- actinobacteria_raw_abundance_distance[,-c(1)]
actinobacteria_raw_abundance_distance <- column_to_rownames(actinobacteria_raw_abundance_distance, "otu_id")
actinobacteria_raw_abundance_distance <- as.matrix(t(actinobacteria_raw_abundance_distance))

bacteroidetes_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Bacteroidetes")

bacteroidetes_raw_abundance_distance <- bacteroidetes_raw_abundance_distance[,-c(1)]
bacteroidetes_raw_abundance_distance <- column_to_rownames(bacteroidetes_raw_abundance_distance, "otu_id")
bacteroidetes_raw_abundance_distance <- as.matrix(t(bacteroidetes_raw_abundance_distance))

chloroflexi_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Chloroflexi")

chloroflexi_raw_abundance_distance <- chloroflexi_raw_abundance_distance[,-c(1)]
chloroflexi_raw_abundance_distance <- column_to_rownames(chloroflexi_raw_abundance_distance, "otu_id")
chloroflexi_raw_abundance_distance <- as.matrix(t(chloroflexi_raw_abundance_distance))

firmicutes_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Firmicutes")

firmicutes_raw_abundance_distance <- firmicutes_raw_abundance_distance[,-c(1)]
firmicutes_raw_abundance_distance <- column_to_rownames(firmicutes_raw_abundance_distance, "otu_id")
firmicutes_raw_abundance_distance <- as.matrix(t(firmicutes_raw_abundance_distance))

gemmatimonadetes_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Gemmatimonadetes")

gemmatimonadetes_raw_abundance_distance <- gemmatimonadetes_raw_abundance_distance[,-c(1)]
gemmatimonadetes_raw_abundance_distance <- column_to_rownames(gemmatimonadetes_raw_abundance_distance, "otu_id")
gemmatimonadetes_raw_abundance_distance <- as.matrix(t(gemmatimonadetes_raw_abundance_distance))

nitrospirae_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Nitrospirae")

nitrospirae_raw_abundance_distance <- nitrospirae_raw_abundance_distance[,-c(1)]
nitrospirae_raw_abundance_distance <- column_to_rownames(nitrospirae_raw_abundance_distance, "otu_id")
nitrospirae_raw_abundance_distance <- as.matrix(t(nitrospirae_raw_abundance_distance))

planctomycetes_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Planctomycetes")

planctomycetes_raw_abundance_distance <- planctomycetes_raw_abundance_distance[,-c(1)]
planctomycetes_raw_abundance_distance <- column_to_rownames(planctomycetes_raw_abundance_distance, "otu_id")
planctomycetes_raw_abundance_distance <- as.matrix(t(planctomycetes_raw_abundance_distance))

proteobacteria_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Proteobacteria")

proteobacteria_raw_abundance_distance <- proteobacteria_raw_abundance_distance[,-c(1)]
proteobacteria_raw_abundance_distance <- column_to_rownames(proteobacteria_raw_abundance_distance, "otu_id")
proteobacteria_raw_abundance_distance <- as.matrix(t(proteobacteria_raw_abundance_distance))

verrucomicrobia_raw_abundance_distance <- soil_data_phylum_top10_raw_abundance_general %>%
  filter(phylum == "Verrucomicrobia")

verrucomicrobia_raw_abundance_distance <- verrucomicrobia_raw_abundance_distance[,-c(1)]
verrucomicrobia_raw_abundance_distance <- column_to_rownames(verrucomicrobia_raw_abundance_distance, "otu_id")
verrucomicrobia_raw_abundance_distance <- as.matrix(t(verrucomicrobia_raw_abundance_distance))
```

## Transform raw abundance data by total abundance per site, and then produce the distacne matrix for each phylum
```{r}
#Transform the data so that each raw abundance value is corrected for the total abundance of each site
acidobacteria_dist_transform <- decostand(acidobacteria_raw_abundance_distance, "total")
actinobacteria_dist_transform <- decostand(actinobacteria_raw_abundance_distance, "total")
bacteroidetes_dist_transform <- decostand(bacteroidetes_raw_abundance_distance, "total")
chloroflexi_dist_transform <- decostand(chloroflexi_raw_abundance_distance, "total")
firmicutes_dist_transform <- decostand(firmicutes_raw_abundance_distance, "total")
gemmatimonadetes_dist_transform <- decostand(gemmatimonadetes_raw_abundance_distance, "total")
nitrospirae_dist_transform <- decostand(nitrospirae_raw_abundance_distance, "total")
planctomycetes_dist_transform <- decostand(planctomycetes_raw_abundance_distance, "total")
proteobacteria_dist_transform <- decostand(proteobacteria_raw_abundance_distance, "total")
verrucomicrobia_dist_transform <- decostand(verrucomicrobia_raw_abundance_distance, "total")

#Create distance matrix for each phylum using Bray-Curtis distance
acidobacteria_dist <- vegdist(acidobacteria_dist_transform, "bray")
actinobacteria_dist <- vegdist(actinobacteria_dist_transform, "bray")
bacteroidetes_dist <- vegdist(bacteroidetes_dist_transform, "bray")
chloroflexi_dist <- vegdist(chloroflexi_dist_transform, "bray")
firmicutes_dist <- vegdist(firmicutes_dist_transform, "bray")
gemmatimonadetes_dist <- vegdist(gemmatimonadetes_dist_transform, "bray")
nitrospirae_dist<- vegdist(nitrospirae_dist_transform, "bray")
planctomycetes_dist <- vegdist(planctomycetes_dist_transform, "bray")
proteobacteria_dist <- vegdist(proteobacteria_dist_transform, "bray")
verrucomicrobia_dist <- vegdist(verrucomicrobia_dist_transform, "bray")
```

## Run NMDS Models
```{r}
#Run NMDS on all phyla
acidobacteria_nmds <- metaMDS(acidobacteria_dist, autotransform = F, expand = F, trace = F)
actinobacteria_nmds <- metaMDS(actinobacteria_dist, autotransform = F, expand = F, trace = F)
bacteroidetes_nmds <- metaMDS(bacteroidetes_dist, autotransform = F, expand = F, trace = F)
chloroflexi_nmds <- metaMDS(chloroflexi_dist, autotransform = F, expand = F, trace = F)
firmicutes_nmds <- metaMDS(firmicutes_dist, autotransform = F, expand = F, trace = F)
gemmatimonadetes_nmds <- metaMDS(gemmatimonadetes_dist, autotransform = F, expand = F, trace = F)
nitrospirae_nmds <- metaMDS(nitrospirae_dist, autotransform = F, expand = F, trace = F)
planctomycetes_nmds <- metaMDS(planctomycetes_dist, autotransform = F, expand = F, trace = F)
proteobacteria_nmds <- metaMDS(proteobacteria_dist, autotransform = F, expand = F, trace = F)
verrucomicrobia_nmds <- metaMDS(verrucomicrobia_dist, autotransform = F, expand = F, trace = F)
```

## Extract NMDS axes and covert into usable tables for plotting
```{R}
#Extract "NMDS Axes" for plotting
acidobacteria_scores <- as.data.frame(scores(acidobacteria_nmds))
actinobacteria_scores <- as.data.frame(scores(actinobacteria_nmds))
bacteroidetes_scores <- as.data.frame(scores(bacteroidetes_nmds))
chloroflexi_scores <- as.data.frame(scores(chloroflexi_nmds))
firmicutes_scores <- as.data.frame(scores(firmicutes_nmds))
gemmatimonadetes_scores <- as.data.frame(scores(gemmatimonadetes_nmds))
nitrospirae_scores <- as.data.frame(scores(nitrospirae_nmds))
planctomycetes_scores <- as.data.frame(scores(planctomycetes_nmds))
proteobacteria_scores <- as.data.frame(scores(proteobacteria_nmds))
verrucomicrobia_scores <- as.data.frame(scores(verrucomicrobia_nmds))

#Turn Site into a viable column so it can be merged with the metadata
acidobacteria_scores <- acidobacteria_scores %>% rownames_to_column("Site")
actinobacteria_scores <- actinobacteria_scores %>% rownames_to_column("Site")
bacteroidetes_scores <- bacteroidetes_scores %>% rownames_to_column("Site")
chloroflexi_scores <- chloroflexi_scores %>% rownames_to_column("Site")
firmicutes_scores <- firmicutes_scores %>% rownames_to_column("Site")
gemmatimonadetes_scores <- gemmatimonadetes_scores %>% rownames_to_column("Site")
nitrospirae_scores <- nitrospirae_scores %>% rownames_to_column("Site")
planctomycetes_scores <- planctomycetes_scores %>% rownames_to_column("Site")
proteobacteria_scores <- proteobacteria_scores %>% rownames_to_column("Site")
verrucomicrobia_scores <- verrucomicrobia_scores %>% rownames_to_column("Site")

#Merge each score dataset with metadata
acidobacteria_scores <- acidobacteria_scores %>% left_join(site_metadata, "Site")
actinobacteria_scores <- actinobacteria_scores %>% left_join(site_metadata, "Site")
bacteroidetes_scores <- bacteroidetes_scores %>% left_join(site_metadata, "Site")
chloroflexi_scores <- chloroflexi_scores %>% left_join(site_metadata, "Site")
firmicutes_scores <- firmicutes_scores %>% left_join(site_metadata, "Site")
gemmatimonadetes_scores <- gemmatimonadetes_scores %>% left_join(site_metadata, "Site")
nitrospirae_scores <- nitrospirae_scores %>% left_join(site_metadata, "Site")
planctomycetes_scores <- planctomycetes_scores %>% left_join(site_metadata, "Site")
proteobacteria_scores <- proteobacteria_scores %>% left_join(site_metadata, "Site")
verrucomicrobia_scores <- verrucomicrobia_scores %>% left_join(site_metadata, "Site")

#Merge each score dataset with metadata
acidobacteria_scores <- acidobacteria_scores %>% mutate(Site = factor(Site))
actinobacteria_scores <- actinobacteria_scores %>% mutate(Site = factor(Site))
bacteroidetes_scores <- bacteroidetes_scores %>% mutate(Site = factor(Site))
chloroflexi_scores <- chloroflexi_scores %>% mutate(Site = factor(Site))
firmicutes_scores <- firmicutes_scores %>% mutate(Site = factor(Site))
gemmatimonadetes_scores <- gemmatimonadetes_scores %>% mutate(Site = factor(Site))
nitrospirae_scores <- nitrospirae_scores %>% mutate(Site = factor(Site))
planctomycetes_scores <- planctomycetes_scores %>% mutate(Site = factor(Site))
proteobacteria_scores <- proteobacteria_scores %>% mutate(Site = factor(Site))
verrucomicrobia_scores <- verrucomicrobia_scores %>% mutate(Site = factor(Site))
```

## Plot NMDS 
```{r}
ggplot(acidobacteria_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(actinobacteria_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(bacteroidetes_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(chloroflexi_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(firmicutes_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(gemmatimonadetes_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(nitrospirae_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(planctomycetes_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(proteobacteria_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()

ggplot(verrucomicrobia_scores, aes(x = NMDS1, y = NMDS2, color = Location)) +
  geom_point()
```


## Get separate datasets for each phylum based on raw data
```{r}
###All datasets below are for individual phyla (top 10) and show raw abundance data
acidobacteria_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Acidobacteria")

actinobacteria_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Actinobacteria")

bacteroidetes_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Bacteroidetes")

chloroflexi_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Chloroflexi")

firmicutes_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Firmicutes")

gemmatimonadetes_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Gemmatimonadetes")

nitrospirae_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Nitrospirae")

planctomycetes_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Planctomycetes")

proteobacteria_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Proteobacteria")

verrucomicrobia_raw_abundance <- phy_10_raw_abundance %>%
  filter(phylum == "Verrucomicrobia")
```



## Visualize and run general linear mixed effect models (with location as a random effect in each model) for each phylum to assess differences among different orchid LOCATION treatments

## ACIDOBACTERIA
```{r}
#Plot raw abundance at different distances
ggplot(acidobacteria_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
acidobacteria_glmer <- glmer.nb(raw_abundance ~ Location + (1|Near_Far), data = acidobacteria_raw_abundance)
```

## ACTINOBACTERIA
```{r}
#Plot raw abundance at different distances
ggplot(actinobacteria_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
actinobacteria_glmer <- glmer.nb(raw_abundance ~ Location + (1|Near_Far), data = actinobacteria_raw_abundance)
```

## BACTEROIDETES
```{r}
#Plot raw abundance at different distances
ggplot(bacteroidetes_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
bacteroidetes_glmer <- glmer.nb(raw_abundance ~ Location + (1|Near_Far), data = bacteroidetes_raw_abundance)
```

## CHLOROFLEXI
```{r}
#Plot raw abundance at different distances
ggplot(chloroflexi_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
chloroflexi_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = chloroflexi_raw_abundance)
```

## FIRMICUTES
```{r}
#Plot raw abundance at different distances
ggplot(firmicutes_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
firmicutes_glmer <- glmer.nb(raw_abundance ~ Location + (1|Near_Far), data = firmicutes_raw_abundance)
```

## GEMMATIMONADETES
```{r}
#Plot raw abundance at different distances
ggplot(gemmatimonadetes_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
gemmatimonadetes_glmer <- glmer.nb(raw_abundance ~ Location + (1|Near_Far), data = gemmatimonadetes_raw_abundance)
```

## NITROSPIRAE
```{r}
#Plot raw abundance at different distances
ggplot(nitrospirae_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
nitrospirae_glmer <- glmer.nb(raw_abundance ~ Location + (1|Near_Far), data = nitrospirae_raw_abundance)
```

## PLANCTOMYCETES
```{r}
#Plot raw abundance at different distances
ggplot(planctomycetes_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
planctomycetes_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = planctomycetes_raw_abundance)
```

## PROTEOBACTERIA
```{r}
#Plot raw abundance at different distances
ggplot(proteobacteria_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
proteobacteria_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = proteobacteria_raw_abundance)
```

## VERRUCOMICROBIA
```{r}
#Plot raw abundance at different distances
ggplot(verrucomicrobia_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(position = position_jitter(width = 0.2), alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
verruco_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = verrucomicrobia_raw_abundance)
```

## All model summary outputs
```{r}
summary(acidobacteria_glmer) #SIGNIFICANT
summary(actinobacteria_glmer)
summary(bacteroidetes_glmer) #SIGNIFICANT
summary(chloroflexi_glmer)
summary(firmicutes_glmer) #SIGNIFICANT
summary(gemmatimonadetes_glmer) #SIGNIFICANT
summary(nitrospirae_glmer) #SIGNIFICANT
summary(planctomycetes_glmer)
summary(proteobacteria_glmer)
summary(verruco_glmer)

#chloroflexi was significant, so I am going to test for pairwise differences between treatments using a post-hoc Tukey test

##Post-hoc Test for chloroflexi_glmer (summary output above)

post_acidobacteria <- glht(acidobacteria_glmer, mcp(Location = "Tukey"))
summary(post_acidobacteria) 

post_bacteroidetes <- glht(bacteroidetes_glmer, mcp(Location = "Tukey"))
summary(post_bacteroidetes) 

post_firmicutes <- glht(firmicutes_glmer, mcp(Location = "Tukey"))
summary(post_firmicutes) 

post_gemmatimonadetes <- glht(gemmatimonadetes_glmer, mcp(Location = "Tukey"))
summary(post_gemmatimonadetes) 

post_nitrospirae <- glht(nitrospirae_glmer, mcp(Location = "Tukey"))
summary(post_nitrospirae) 
```

## Plot Location results
```{r}
acidobacteria_location_plot <- ggplot(acidobacteria_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6, fill = "grey") +
  geom_point(alpha = 0.8, show.legend = F) +
  labs(x = "Location", y = "Raw Abundance") +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks.x = element_blank())

bacteroidetes_location_plot <- ggplot(bacteroidetes_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6, fill = "grey") +
  geom_point(alpha = 0.8, show.legend = F) +
  labs(x = "Location", y = "Raw Abundance") +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks.x = element_blank())

gemmatimonadetes_location_plot <- ggplot(gemmatimonadetes_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6, fill = "grey") +
  geom_point(alpha = 0.8, show.legend = F) +
  labs(x = "Location", y = "Raw Abundance") +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks.x = element_blank())

firmicutes_location_plot <- ggplot(firmicutes_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6, fill = "grey") +
  geom_point(alpha = 0.8, show.legend = F) +
  labs(x = "Location", y = "Raw Abundance") +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks.x = element_blank())

nitrospirae_location_plot <- ggplot(nitrospirae_raw_abundance, aes(x = Location, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6, fill = "grey") +
  geom_point(alpha = 0.8, show.legend = F) +
  labs(x = "Location", y = "Raw Abundance") +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 15))


pdf("/Users/joegunn/Desktop/Grad_School_Stuff/Research/Projects/Soil_Microbiome/visualization/phylum_comparison_figures/phylum_location_plots.pdf", width = 9, height = 18)

plot_grid(acidobacteria_location_plot, bacteroidetes_location_plot, gemmatimonadetes_location_plot, firmicutes_location_plot, nitrospirae_location_plot, nrow = 5, labels = c("a","b","c","d","e"), label_size = 25)

dev.off()

##This figure will be altered to add the post-hoc significance values from the package multcomp.
  
```

```{r}
###All datasets below are for individual phyla (top 10) and show raw abundance data
LT_rel_abundance <- phy_10_rel_abundance %>%
  filter(Location == "LT") 

NE_rel_abundance <- phy_10_rel_abundance %>%
  filter(Location == "NE")

SWALE_rel_abundance <- phy_10_rel_abundance %>%
  filter(Location == "Swale")

TARKIO_rel_abundance <- phy_10_rel_abundance %>%
  filter(Location == "Tarkio")

HELTON_rel_abundance <- phy_10_rel_abundance %>%
  filter(Location == "Helton")

PAWNEE_rel_abundance <- phy_10_rel_abundance %>%
  filter(Location == "Pawnee")
```

## Visualize Top 10 phyla relative abundance in each location
```{r}
LT_rel_abundance$phylum <- factor(LT_rel_abundance$phylum, levels = c("Verrucomicrobia","Proteobacteria","Actinobacteria","Acidobacteria","Chloroflexi","Planctomycetes", "Bacteroidetes","Firmicutes","Nitrospirae","Gemmatimonadetes")) #reorder levels - otherwise it will automatically list aplphabetically

NE_rel_abundance$phylum <- factor(NE_rel_abundance$phylum, levels = c("Verrucomicrobia","Proteobacteria","Actinobacteria","Acidobacteria","Chloroflexi","Planctomycetes", "Bacteroidetes","Firmicutes","Nitrospirae","Gemmatimonadetes")) #reorder levels - otherwise it will automatically list aplphabetically

SWALE_rel_abundance$phylum <- factor(SWALE_rel_abundance$phylum, levels = c("Verrucomicrobia","Proteobacteria","Actinobacteria","Acidobacteria","Chloroflexi","Planctomycetes", "Bacteroidetes","Firmicutes","Nitrospirae","Gemmatimonadetes")) #reorder levels - otherwise it will automatically list aplphabetically

TARKIO_rel_abundance$phylum <- factor(TARKIO_rel_abundance$phylum, levels = c("Verrucomicrobia","Proteobacteria","Actinobacteria","Acidobacteria","Chloroflexi","Planctomycetes", "Bacteroidetes","Firmicutes","Nitrospirae","Gemmatimonadetes")) #reorder levels - otherwise it will automatically list aplphabetically

HELTON_rel_abundance$phylum <- factor(HELTON_rel_abundance$phylum, levels = c("Verrucomicrobia","Proteobacteria","Actinobacteria","Acidobacteria","Chloroflexi","Planctomycetes", "Bacteroidetes","Firmicutes","Nitrospirae","Gemmatimonadetes")) #reorder levels - otherwise it will automatically list aplphabetically

PAWNEE_rel_abundance$phylum <- factor(PAWNEE_rel_abundance$phylum, levels = c("Verrucomicrobia","Proteobacteria","Actinobacteria","Acidobacteria","Chloroflexi","Planctomycetes", "Bacteroidetes","Firmicutes","Nitrospirae","Gemmatimonadetes")) #reorder levels - otherwise it will automatically list aplphabetically

HELTON_phylum_rel_abundance_plot <- ggplot(HELTON_rel_abundance, aes(x = phylum, y = relative_abundance)) + 
  geom_boxplot(show.legend = F, outlier.shape = NA, alpha = 0.6, fill = "#1B9E77") + 
  geom_point(position = position_jitter(width = 0.2), alpha = 0.6, color = "black", fill = "#1B9E77", pch = 21, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "Phylum", y = "Relative Abundance") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 0.95)) +
  theme(axis.ticks.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank())

LT_phylum_rel_abundance_plot <- ggplot(LT_rel_abundance, aes(x = phylum, y = relative_abundance)) + 
  geom_boxplot(show.legend = F, outlier.shape = NA, alpha = 0.6, fill = "#D95F02") + 
  geom_point(position = position_jitter(width = 0.2), alpha = 0.6, color = "black", fill = "#D95F02", pch = 21, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "Phylum", y = "Relative Abundance") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 0.95)) +
  theme(axis.ticks.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank())

NE_phylum_rel_abundance_plot <- ggplot(NE_rel_abundance, aes(x = phylum, y = relative_abundance)) + 
  geom_boxplot(show.legend = F, outlier.shape = NA, alpha = 0.6, fill = "#7570B3") + 
  geom_point(position = position_jitter(width = 0.2), alpha = 0.6, color = "black", fill = "#7570B3", pch = 21, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "Phylum", y = "Relative Abundance") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 0.95)) +
  theme(axis.ticks.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank())

PAWNEE_phylum_rel_abundance_plot <- ggplot(PAWNEE_rel_abundance, aes(x = phylum, y = relative_abundance)) + 
  geom_boxplot(show.legend = F, outlier.shape = NA, alpha = 0.6, fill = "#E7298A") + 
  geom_point(position = position_jitter(width = 0.2), alpha = 0.6, color = "black", fill = "#E7298A", pch = 21, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "Phylum", y = "Relative Abundance") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 0.95)) +
  theme(axis.ticks.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank())

SWALE_phylum_rel_abundance_plot <- ggplot(SWALE_rel_abundance, aes(x = phylum, y = relative_abundance)) + 
  geom_boxplot(show.legend = F, outlier.shape = NA, alpha = 0.6, fill = "#66A61E") + 
  geom_point(position = position_jitter(width = 0.2), alpha = 0.6, color = "black", fill = "#66A61E", pch = 21, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "Phylum", y = "Relative Abundance") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 0.95)) +
  theme(axis.ticks.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank())

TARKIO_phylum_rel_abundance_plot <- ggplot(TARKIO_rel_abundance, aes(x = phylum, y = relative_abundance)) + 
  geom_boxplot(show.legend = F, outlier.shape = NA, alpha = 0.6, fill = "#E6AB02") + 
  geom_point(position = position_jitter(width = 0.2), alpha = 0.6, color = "black", fill = "#E6AB02", pch = 21, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "Phylum", y = "Relative Abundance") +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 0.95)) +
  theme(axis.ticks.x = element_blank())  +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_blank())

pdf("/Users/joegunn/Desktop/Grad_School_Stuff/Research/Projects/Soil_Microbiome/visualization/phylum_comparison_figures/phylum_rel_abundance_by_location.pdf", width = 8, height = 20)

plot_grid(HELTON_phylum_rel_abundance_plot, LT_phylum_rel_abundance_plot, NE_phylum_rel_abundance_plot, PAWNEE_phylum_rel_abundance_plot, SWALE_phylum_rel_abundance_plot, TARKIO_phylum_rel_abundance_plot, ncol = 1, labels = c("a","b","c","d","e","f"), label_size = 25)

dev.off()
```

## Filter each one of the above datasets to only include Near and No orchid treatments - this will try to get at whether there are any phyla that are only represented near to orchids
```{r}
acidobacteria_orchids <- acidobacteria_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

actinobacteria_orchids <- actinobacteria_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

bacteroidetes_orchids <- bacteroidetes_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

chloroflexi_orchids <- chloroflexi_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

firmicutes_orchids <- firmicutes_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

gemmatimonadetes_orchids <- gemmatimonadetes_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

nitrospirae_orchids <- nitrospirae_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

planctomycetes_orchids <- planctomycetes_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

proteobacteria_orchids <- proteobacteria_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

verrucomicrobia_orchids <- verrucomicrobia_raw_abundance %>%
  filter(Near_Far == "Near" | Near_Far == "NO" | Near_Far == "NOE")

```

## Visualize and run general linear mixed effect models (with location as a random effect in each model) for each phylum to assess differences among different orchid DISTANCE treatments

## ACIDOBACTERIA
```{r}
#Plot raw abundance at different distances
ggplot(acidobacteria_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
acidobacteria_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = acidobacteria_raw_abundance)
```

## ACTINOBACTERIA
```{r}
#Plot raw abundance at different distances
ggplot(actinobacteria_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
actinobacteria_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = actinobacteria_raw_abundance)
```

## BACTEROIDETES
```{r}
#Plot raw abundance at different distances
ggplot(bacteroidetes_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
bacteroidetes_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = bacteroidetes_raw_abundance)
```

## CHLOROFLEXI
```{r}
#Plot raw abundance at different distances
ggplot(chloroflexi_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
chloroflexi_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = chloroflexi_raw_abundance)
```

## FIRMICUTES
```{r}
#Plot raw abundance at different distances
ggplot(firmicutes_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
firmicutes_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = firmicutes_raw_abundance)
```

## GEMMATIMONADETES
```{r}
#Plot raw abundance at different distances
ggplot(gemmatimonadetes_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
gemmatimonadetes_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = gemmatimonadetes_raw_abundance)
```

## NITROSPIRAE
```{r}
#Plot raw abundance at different distances
ggplot(nitrospirae_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
nitrospirae_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = nitrospirae_raw_abundance)
```

## PLANCTOMYCETES
```{r}
#Plot raw abundance at different distances
ggplot(planctomycetes_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
planctomycetes_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = planctomycetes_raw_abundance)
```

## PROTEOBACTERIA
```{r}
#Plot raw abundance at different distances
ggplot(proteobacteria_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
proteobacteria_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = proteobacteria_raw_abundance)
```

## VERRUCOMICROBIA
```{r}
#Plot raw abundance at different distances
ggplot(verrucomicrobia_raw_abundance, aes(x = Near_Far, y = raw_abundance)) +
  geom_boxplot(show.legend = T, outlier.shape = NA, alpha = 0.6) +
  geom_point(position = position_jitter(width = 0.2), alpha = 0.8, show.legend = F)

#Run general linear mixed effect model with negative binomial distribution (as inferred from ORCHIDS_distribution_analysis.rmd)
verruco_glmer <- glmer.nb(raw_abundance ~ Near_Far + (1|Location), data = verrucomicrobia_raw_abundance)
```

## All model summary outputs
```{r}
#summary(acidobacteria_glmer) #For some reason this model would not run, but the graph makes it clear that there is no signficant difference among distances
summary(actinobacteria_glmer)
summary(bacteroidetes_glmer)
summary(chloroflexi_glmer)
summary(firmicutes_glmer)
summary(gemmatimonadetes_glmer)
summary(nitrospirae_glmer)
summary(planctomycetes_glmer)
summary(proteobacteria_glmer)
summary(verruco_glmer)

post_chloroflexi <- glht(chloroflexi_glmer, mcp(Near_Far = "Tukey"))
summary(post_chloroflexi) #For some reason there are no significant differences here, even though the overall model was significant
```


## Visualize possible distribution fits
```{r}
#Get coefficients for normal distribution
mu_mle_norm <- coef(fit_norm)[[1]]
sd_mle_norm <- coef(fit_norm)[[2]]

#Get coefficients for poisson distribution
lambda_mle_pois <- coef(fit_pois)[[1]]

#Get coefficients for negative binomial distribution
mu_mle_nbinom <- coef(fit_nbinom)[[1]]
size_mle_nbinom <- coef(fit_nbinom)[[2]]

#Get coefficients for weibull distribution
shape_mle_weibull <- coef(fit_weibull)[[1]]

par(mfrow=c(1,1))
plot(density(phy_10_raw_abundance$raw_abundance), xlab = 'Raw Abundance', lwd = 4, ylim = c(0, .001), xlim = c(0, 30000), main = "Fit of Distributions for Raw Abundance Data")
curve(dpois(x, lambda = lambda_mle_pois), 0, 5000, col = 'blue', lwd = 2, add = T)
curve(dnbinom(x, mu = mu_mle_nbinom, size = size_mle_nbinom), 0, 5000, col = 'orange', lwd = 4, lty = 2, add = T)
curve(dnorm(x, mean = mu_mle_norm, sd = sd_mle_norm), -10, 5000, col = 'green', lwd = 2, add = T)
curve(dweibull(x, shape = shape_mle_weibull), 0, 5000, col = 'red', lwd = 2, add = T)
legend(6, 0.4, legend = c('data dist', 'poisson dist', 'neg. binomial dist', 'normal dist', 'weibull dist'), lty = 1, col = c('black','blue','orange', "green", "red"), lwd = c(4,2,2,2,2))

```
