---
title: "ORCHIDS_ORDER_COMPARISON"
author: "Joe Gunn"
date: "5/21/2020"
output: html_document
---


## Libraries needed for analysis
```{r libraries, echo=FALSE}
library(readxl)
library(cowplot)
library(tidyverse)
library(vegan)
library(nlme)
library(lme4)
library(factoextra)
library(viridis)
library(bbmle)
library(car)
```

## Metadata
```{r}
#Read in site-level metadata and clean
site_metadata <- read_excel("../../data/metadata/site_metadata.xlsx")

site_metadata <- site_metadata %>%
  mutate(Site = factor(Site), Near_Far = factor(Near_Far), Location = factor(Location), loc_abbr = factor(loc_abbr)) %>%
  mutate(present_absent = ifelse(as.factor(Near_Far) == "NO" | as.factor(Near_Far) == "NOE", "Absent", "Present"))

#Read in taxonomic level metadata and clean
tax_metadata <- read_excel("../../data/metadata/tax_metadata.xlsx")

tax_metadata <- tax_metadata %>%
  mutate(kingdom = factor(kingdom),
         phylum = factor(phylum), 
         class = factor(class), 
         order = factor(order), 
         family = factor(family), 
         genus = factor(genus), 
         species = factor(species))
```

## Read-in raw data
```{r}
#Read in OTU raw abundance data. This data is organized as a matrix of individual microbial OTUS (rows) and sampling sites (columns). 
soil_raw <- read_excel("../../data/excel_data/raw_data.xlsx")

#Clean table
colnames(soil_raw)[1] <- "otu_id"

soil_raw <- soil_raw %>%
  mutate(otu_id = as.character(otu_id)) %>%
  mutate(otu_id = factor(otu_id))

#Summarize raw microbial OTU data
#Get number of OTUs
otu_count <- soil_raw %>%
  nrow() #84,960 OTUs

#Get number of sites
site_count <- soil_raw %>%
  ncol() #24 sites (excluding otu id)

read_count <- soil_raw[,-c(1)] %>%
  colSums() #54,631 reads per sample - THE DATA IS ALREADY RAREFIED!


#combine taxonomy data (tax_metadatat) and raw otu data (soil_raw)
soil_all <- cbind(tax_metadata, soil_raw)
```

## Get Relative Abundance Data at the Phylum Level in a gathered dataset - this was produced in ORCHIDS_tax_confidence.rmd
```{r}
soil_data_phylum_raw <- soil_all[,-c(1,3:7)]

soil_data_phylum_rel <- soil_data_phylum_raw %>%
  mutate(LT0116S = LT0116S/54631,
         LT0216S = LT0216S/54631,
         NE0616S = NE0616S/54631,
         P0116S = P0116S/54631,
         P0216S = P0216S/54631,
         P0316S = P0316S/54631,
         H0316S = H0316S/54631,
         H0216S = H0216S/54631,
         H0116S = H0116S/54631,
         LT0516S = LT0516S/54631,
         T0416S = T0416S/54631,
         NE0216S = NE0216S/54631,
         NE0416S = NE0416S/54631,
         S0216S = S0216S/54631,
         LT0416S = LT0416S/54631,
         NE0516S = NE0516S/54631,
         S0116S = S0116S/54631,
         LT0316S = LT0316S/54631,
         NE0316S = NE0316S/54631,
         NE0116S = NE0116S/54631,
         T0316S = T0316S/54631,
         NE0816S = NE0816S/54631,
         NE0716S = NE0716S/54631,
         T0116S = T0116S/54631)

#Full Dataset with raw and relative abundance data
soil_data_phylum_top10_raw_abundance_general <- soil_data_phylum_raw %>% 
  filter(phylum == "Verrucomicrobia" | phylum == "Proteobacteria" | phylum == "Actinobacteria" | phylum == "Acidobacteria" | phylum == "Chloroflexi" | phylum == "Planctomycetes" | phylum == "Bacteroidetes" | phylum == "Firmicutes" | phylum == "Nitrospirae" | phylum == "Gemmatimonadetes")
  
soil_data_phylum_top10_relative_abundance_general <- soil_data_phylum_rel %>% 
  filter(phylum == "Verrucomicrobia" | phylum == "Proteobacteria" | phylum == "Actinobacteria" | phylum == "Acidobacteria" | phylum == "Chloroflexi" | phylum == "Planctomycetes" | phylum == "Bacteroidetes" | phylum == "Firmicutes" | phylum == "Nitrospirae" | phylum == "Gemmatimonadetes")
```

## Get raw abundance and relative abundance datasets that are ready to use for all analyses
```{r}
tax_metadata_for_top_10_phyla <- soil_all[,-c(9:32)] %>%
  filter(phylum == "Verrucomicrobia" | phylum == "Proteobacteria" | phylum == "Actinobacteria" | phylum == "Acidobacteria" | phylum == "Chloroflexi" | phylum == "Planctomycetes" | phylum == "Bacteroidetes" | phylum == "Firmicutes" | phylum == "Nitrospirae" | phylum == "Gemmatimonadetes")

soil_data_phylum_top10_raw_abundance_withTAX <- cbind(tax_metadata_for_top_10_phyla, soil_data_phylum_top10_raw_abundance_general) 
soil_data_phylum_top10_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_withTAX[,-c(9:10)]
soil_data_phylum_top10_relative_abundance_withTAX <- cbind(tax_metadata_for_top_10_phyla, soil_data_phylum_top10_relative_abundance_general) 
soil_data_phylum_top10_relative_abundance_withTAX <- soil_data_phylum_top10_relative_abundance_withTAX[,-c(9:10)]

soil_data_phylum_top10_raw_abundance <- column_to_rownames(soil_data_phylum_top10_raw_abundance_general, "otu_id")
soil_data_phylum_top10_relative_abundance <- column_to_rownames(soil_data_phylum_top10_relative_abundance_general, "otu_id")
```

## Get separate datasets without SWALE
```{r}
soil_data_phylum_top10_raw_abundance_noSWALE <- soil_data_phylum_top10_raw_abundance[,-c(15,18)]
soil_data_phylum_top10_relative_abundance_noSWALE <- soil_data_phylum_top10_relative_abundance[,-c(15,18)]

soil_data_phylum_top10_raw_abundance_noSWALE_withTAX <- cbind(tax_metadata_for_top_10_phyla, soil_data_phylum_top10_raw_abundance_noSWALE) 
soil_data_phylum_top10_raw_abundance_noSWALE_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX[,-c(8:9)]
```

## Get top 10 phyla with all other taxonomic data
```{r}
#Here I am separating the data into separate phyla, but I am keeping all other taxonomic information as well so that I can eventually extract orders 
acidobacteria_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Acidobacteria")

actinobacteria_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Actinobacteria")

bacteroidetes_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Bacteroidetes")

chloroflexi_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Chloroflexi")

firmicutes_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Firmicutes")

gemmatimonadetes_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Gemmatimonadetes")

nitrospirae_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Nitrospirae")

planctomycetes_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Planctomycetes")

proteobacteria_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Proteobacteria")

verrucomicrobia_raw_abundance_withTAX <- soil_data_phylum_top10_raw_abundance_noSWALE_withTAX %>%
  filter(phylum == "Verrucomicrobia")
```

## Get the significant orders within each of the top 5 phyla from my absence/presence analysis (ORCHIDS_near_vs_no_orchids_comparison.rmd)
## ACIDOBACTERIA
```{r}
#Exclude all taxonomic metadata except for orders
acidobacteria_raw_abundance_withTAX_orders <- acidobacteria_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 10,629 total OTUs
acidobacteria_raw_abundance_withTAX_orders <- acidobacteria_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
acidobacteria_orders_sums <- acidobacteria_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
acidobacteria_orders_sum_inverted <- acidobacteria_orders_sums %>% column_to_rownames("order")
acidobacteria_orders_sum_inverted <- data.frame(t(acidobacteria_orders_sum_inverted))
acidobacteria_orders_sum_inverted <- acidobacteria_orders_sum_inverted %>% rownames_to_column("Site")
acidobacteria_orders_sum_inverted <- acidobacteria_orders_sum_inverted %>%
  mutate(Site = factor(Site))
acidobacteria_orders_sum_inverted_withTAX <- acidobacteria_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

acidobacteria_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], acidobacteria_orders_sum_inverted_withTAX)
acidobacteria_orders_sum_inverted_withTAX <- acidobacteria_orders_sum_inverted_withTAX[,-c(6,10:11,15,27:30)]

#Orders B110, BPC015, and E29 were removed due to very low read counts across sites

#17 total orders used in analysis

p_vals_acidobacteria_orders <- data.frame("p_val" = numeric(length = 17))

    for (ii in 6:ncol(acidobacteria_orders_sum_inverted_withTAX)) {
  
      col = acidobacteria_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(acidobacteria_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glm_temp <- glm.nb(ff, data = acidobacteria_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glm_temp)$coefficients[2,4] #extracts p value
      p_vals_acidobacteria_orders[ii-5,"p_val"] <- p_value
    }

acidobacteria_pval_fdr <- p.adjust(p_vals_acidobacteria_orders$p_val, method = "fdr")

#No significant orders in this phylum

summary(glm.nb(X11.24 ~ present_absent, data = acidobacteria_orders_sum_inverted_withTAX))
```

## ACTINOBACTERIA
```{r}
#Exclude all taxonomic metadata except for orders
actinobacteria_raw_abundance_withTAX_orders <- actinobacteria_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 10,629 total OTUs
actinobacteria_raw_abundance_withTAX_orders <- actinobacteria_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
actinobacteria_orders_sums <- actinobacteria_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
actinobacteria_orders_sum_inverted <- actinobacteria_orders_sums %>% column_to_rownames("order")
actinobacteria_orders_sum_inverted <- data.frame(t(actinobacteria_orders_sum_inverted))
actinobacteria_orders_sum_inverted <- actinobacteria_orders_sum_inverted %>% rownames_to_column("Site")
actinobacteria_orders_sum_inverted <- actinobacteria_orders_sum_inverted %>%
  mutate(Site = factor(Site))
actinobacteria_orders_sum_inverted_withTAX <- actinobacteria_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

actinobacteria_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], actinobacteria_orders_sum_inverted_withTAX)
actinobacteria_orders_sum_inverted_withTAX <- actinobacteria_orders_sum_inverted_withTAX[,-c(6,10:11,14,16:20)]

#Orders B110, BPC015, and E29 were removed due to very low read counts across sites

#17 total orders used in analysis

p_vals_actinobacteria_orders <- data.frame("p_val" = numeric(length = 6))

    for (ii in 6:ncol(actinobacteria_orders_sum_inverted_withTAX)) {
  
      col = actinobacteria_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(actinobacteria_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glm_temp <- glm.nb(ff, data = actinobacteria_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glm_temp)$coefficients[2,4] #extracts p value
      p_vals_actinobacteria_orders[ii-5,"p_val"] <- p_value
    }

actinobacteria_pval_fdr <- p.adjust(p_vals_actinobacteria_orders$p_val, method = "fdr")

```

## BACTEROIDETES
```{r}
#Exclude all taxonomic metadata except for orders
bacteroidetes_raw_abundance_withTAX_orders <- bacteroidetes_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 10,629 total OTUs
bacteroidetes_raw_abundance_withTAX_orders <- bacteroidetes_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
bacteroidetes_orders_sums <- bacteroidetes_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
bacteroidetes_orders_sum_inverted <- bacteroidetes_orders_sums %>% column_to_rownames("order")
bacteroidetes_orders_sum_inverted <- data.frame(t(bacteroidetes_orders_sum_inverted))
bacteroidetes_orders_sum_inverted <- bacteroidetes_orders_sum_inverted %>% rownames_to_column("Site")
bacteroidetes_orders_sum_inverted <- bacteroidetes_orders_sum_inverted %>%
  mutate(Site = factor(Site))
bacteroidetes_orders_sum_inverted_withTAX <- bacteroidetes_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

bacteroidetes_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], bacteroidetes_orders_sum_inverted_withTAX)
bacteroidetes_orders_sum_inverted_withTAX <- bacteroidetes_orders_sum_inverted_withTAX[,-c(6,7,13:16)]

#Orders B110, BPC015, and E29 were removed due to very low read counts across sites

#17 total orders used in analysis

p_vals_bacteroidetes_orders <- data.frame("p_val" = numeric(length = 5))

    for (ii in 6:ncol(bacteroidetes_orders_sum_inverted_withTAX)) {
  
      col = bacteroidetes_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(bacteroidetes_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glm_temp <- glm.nb(ff, data = bacteroidetes_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glm_temp)$coefficients[2,4] #extracts p value
      p_vals_bacteroidetes_orders[ii-5,"p_val"] <- p_value
    }

bacteroidetes_pval_fdr <- p.adjust(p_vals_bacteroidetes_orders$p_val, method = "fdr")

#No significant orders in this phylum
```

## CHLOROFLEXI
```{r}
#Exclude all taxonomic metadata except for orders
chloroflexi_raw_abundance_withTAX_orders <- chloroflexi_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 3,729 total OTUs
chloroflexi_raw_abundance_withTAX_orders <- chloroflexi_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
chloroflexi_orders_sums <- chloroflexi_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
chloroflexi_orders_sum_inverted <- chloroflexi_orders_sums %>% column_to_rownames("order")
chloroflexi_orders_sum_inverted <- data.frame(t(chloroflexi_orders_sum_inverted))
chloroflexi_orders_sum_inverted <- chloroflexi_orders_sum_inverted %>% rownames_to_column("Site")
chloroflexi_orders_sum_inverted <- chloroflexi_orders_sum_inverted %>%
  mutate(Site = factor(Site))
chloroflexi_orders_sum_inverted_withTAX <- chloroflexi_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

chloroflexi_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], chloroflexi_orders_sum_inverted_withTAX)
chloroflexi_orders_sum_inverted_withTAX <- chloroflexi_orders_sum_inverted_withTAX[,-c(6,8,14,24,32,36:37,41:44)]

#Orders Thermobaculales, Ardenscatenales, FS117.23B.02, pLW.97, SJA.15, and Sphaerobacterales were removed due to very low read counts across sites

#28 total orders used in analysis

###Ran this with a standard glm, because the mixed effect model would not run - didn't matter - nothing was significant anyway
p_vals_chloroflexi_orders <- data.frame("p_val" = numeric(length = 28))

    for (ii in 6:ncol(chloroflexi_orders_sum_inverted_withTAX)) {
  
      col = chloroflexi_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(chloroflexi_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glmer_temp <- glm.nb(ff, data = chloroflexi_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glmer_temp)$coefficients[2,4] #extracts p value
      p_vals_chloroflexi_orders[ii-5,"p_val"] <- p_value
    }

chloroflexi_pval_fdr <- p.adjust(p_vals_chloroflexi_orders$p_val, method = "fdr")

#No significant orders in this phylum
```

## FIRMICUTES
```{r}
#Exclude all taxonomic metadata except for orders
firmicutes_raw_abundance_withTAX_orders <- firmicutes_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 1,225 total OTUs
firmicutes_raw_abundance_withTAX_orders <- firmicutes_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
firmicutes_orders_sums <- firmicutes_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
firmicutes_orders_sum_inverted <- firmicutes_orders_sums %>% column_to_rownames("order")
firmicutes_orders_sum_inverted <- data.frame(t(firmicutes_orders_sum_inverted))
firmicutes_orders_sum_inverted <- firmicutes_orders_sum_inverted %>% rownames_to_column("Site")
firmicutes_orders_sum_inverted <- firmicutes_orders_sum_inverted %>%
  mutate(Site = factor(Site))
firmicutes_orders_sum_inverted_withTAX <- firmicutes_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

firmicutes_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], firmicutes_orders_sum_inverted_withTAX)
firmicutes_orders_sum_inverted_withTAX <- firmicutes_orders_sum_inverted_withTAX[,-c(6,9:13,15:20)]

#Orders Erysipelotrichales, Halanaerobiales, Lactobacillales, MBA08, Natranaerobiales, Thermoanaerobacterales, and Turicibacterales were removed due to very low read counts across sites

#3 total orders used in analysis

p_vals_firmicutes_orders <- data.frame("p_val" = numeric(length = 3))

    for (ii in 6:ncol(firmicutes_orders_sum_inverted_withTAX)) {
  
      col = firmicutes_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(firmicutes_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glmer_temp <- glm.nb(ff, data = firmicutes_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glmer_temp)$coefficients[2,4] #extracts p value
      p_vals_firmicutes_orders[ii-5,"p_val"] <- p_value
    }

firmicutes_pval_fdr <- p.adjust(p_vals_firmicutes_orders$p_val, method = "fdr")

#No significant orders in this phylum
```

## GEMMATIMONADETES
```{r}
#Exclude all taxonomic metadata except for orders
gemmatimonadetes_raw_abundance_withTAX_orders <- gemmatimonadetes_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 715 total OTUs
gemmatimonadetes_raw_abundance_withTAX_orders <- gemmatimonadetes_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
gemmatimonadetes_orders_sums <- gemmatimonadetes_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
gemmatimonadetes_orders_sum_inverted <- gemmatimonadetes_orders_sums %>% column_to_rownames("order")
gemmatimonadetes_orders_sum_inverted <- data.frame(t(gemmatimonadetes_orders_sum_inverted))
gemmatimonadetes_orders_sum_inverted <- gemmatimonadetes_orders_sum_inverted %>% rownames_to_column("Site")
gemmatimonadetes_orders_sum_inverted <- gemmatimonadetes_orders_sum_inverted %>%
  mutate(Site = factor(Site))
gemmatimonadetes_orders_sum_inverted_withTAX <- gemmatimonadetes_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

gemmatimonadetes_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], gemmatimonadetes_orders_sum_inverted_withTAX)
gemmatimonadetes_orders_sum_inverted_withTAX <- gemmatimonadetes_orders_sum_inverted_withTAX[,-c(6,7,12:15)]

#Order C114 was removed due to very low read counts across sites

#4 total orders used in analysis

p_vals_gemmatimonadetes_orders <- data.frame("p_val" = numeric(length = 4))

    for (ii in 6:ncol(gemmatimonadetes_orders_sum_inverted_withTAX)) {
  
      col = gemmatimonadetes_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(gemmatimonadetes_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glmer_temp <- glm.nb(ff, data = gemmatimonadetes_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glmer_temp)$coefficients[2,4] #extracts p value
      p_vals_gemmatimonadetes_orders[ii-5,"p_val"] <- p_value
    }

gemmatimonadetes_pval_fdr <- p.adjust(p_vals_gemmatimonadetes_orders$p_val, method = "fdr")

#No significant orders in this phylum
```

## NITROSPIRAE
```{r}
#Exclude all taxonomic metadata except for orders
nitrospirae_raw_abundance_withTAX_orders <- nitrospirae_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 1,104 total OTUs
nitrospirae_raw_abundance_withTAX_orders <- nitrospirae_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
nitrospirae_orders_sums <- nitrospirae_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
nitrospirae_orders_sum_inverted <- nitrospirae_orders_sums %>% column_to_rownames("order")
nitrospirae_orders_sum_inverted <- data.frame(t(nitrospirae_orders_sum_inverted))
nitrospirae_orders_sum_inverted <- nitrospirae_orders_sum_inverted %>% rownames_to_column("Site")
nitrospirae_orders_sum_inverted <- nitrospirae_orders_sum_inverted %>%
  mutate(Site = factor(Site))
nitrospirae_orders_sum_inverted_withTAX <- nitrospirae_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

nitrospirae_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], nitrospirae_orders_sum_inverted_withTAX)
nitrospirae_orders_sum_inverted_withTAX <- nitrospirae_orders_sum_inverted_withTAX[,-c(6,8:11)]

#1 total order used in analysis: Nitrospirales

nitrospirae_glmer <- glm.nb(Nitrospirales ~ present_absent, data = nitrospirae_orders_sum_inverted_withTAX)
summary(nitrospirae_glmer)

#No significant orders in this phylum
```

## PLANCTOMYCETES
```{r}
#Exclude all taxonomic metadata except for orders
planctomycetes_raw_abundance_withTAX_orders <- planctomycetes_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 6,764 total OTUs
planctomycetes_raw_abundance_withTAX_orders <- planctomycetes_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
planctomycetes_orders_sums <- planctomycetes_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
planctomycetes_orders_sum_inverted <- planctomycetes_orders_sums %>% column_to_rownames("order")
planctomycetes_orders_sum_inverted <- data.frame(t(planctomycetes_orders_sum_inverted))
planctomycetes_orders_sum_inverted <- planctomycetes_orders_sum_inverted %>% rownames_to_column("Site")
planctomycetes_orders_sum_inverted <- planctomycetes_orders_sum_inverted %>%
  mutate(Site = factor(Site))
planctomycetes_orders_sum_inverted_withTAX <- planctomycetes_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

planctomycetes_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], planctomycetes_orders_sum_inverted_withTAX)
planctomycetes_orders_sum_inverted_withTAX <- planctomycetes_orders_sum_inverted_withTAX[,-c(6,8,10:11,14,18,21,27:29,31:34)]

#Orders AKAU3564, Brocadiales, CCM11a, MSBL9, PeHg47, d113, SHA.43, T8.B82, and Ucm1571 were removed due to very low read counts across sites

#15 total orders used in analysis

p_vals_planctomycetes_orders <- data.frame("p_val" = numeric(length = 15))

    for (ii in 6:ncol(planctomycetes_orders_sum_inverted_withTAX)) {
  
      col = planctomycetes_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(planctomycetes_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glmer_temp <- glm.nb(ff, data = planctomycetes_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glmer_temp)$coefficients[2,4] #extracts p value
      p_vals_planctomycetes_orders[ii-5,"p_val"] <- p_value
    }

planctomycetes_pval_fdr <- p.adjust(p_vals_planctomycetes_orders$p_val, method = "fdr")

#No significant orders in this phylum
```

## PROTEOBACTERIA
```{r}
#Exclude all taxonomic metadata except for orders
proteobacteria_raw_abundance_withTAX_orders <- proteobacteria_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 22,774 total OTUs
proteobacteria_raw_abundance_withTAX_orders <- proteobacteria_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
proteobacteria_orders_sums <- proteobacteria_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
proteobacteria_orders_sum_inverted <- proteobacteria_orders_sums %>% column_to_rownames("order")
proteobacteria_orders_sum_inverted <- data.frame(t(proteobacteria_orders_sum_inverted))
proteobacteria_orders_sum_inverted <- proteobacteria_orders_sum_inverted %>% rownames_to_column("Site")
proteobacteria_orders_sum_inverted <- proteobacteria_orders_sum_inverted %>%
  mutate(Site = factor(Site))
proteobacteria_orders_sum_inverted_withTAX <- proteobacteria_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

proteobacteria_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], proteobacteria_orders_sum_inverted_withTAX)
proteobacteria_orders_sum_inverted_withTAX <- proteobacteria_orders_sum_inverted_withTAX[,-c(6,8:9,11:12,13,15,18,20,24:26,32:34,36:38,40:42,47,51,52,54,56,62:63,67:68,70:71,74:77)]

#36 total orders used in analysis

p_vals_proteobacteria_orders <- data.frame("p_val" = numeric(length = 36))

    for (ii in 6:ncol(proteobacteria_orders_sum_inverted_withTAX)) {
  
      col = proteobacteria_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(proteobacteria_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glmer_temp <- glm.nb(ff, data = proteobacteria_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glmer_temp)$coefficients[2,4] #extracts p value
      p_vals_proteobacteria_orders[ii-5,"p_val"] <- p_value
    }

proteobacteria_pval_fdr <- p.adjust(p_vals_proteobacteria_orders$p_val, method = "fdr")

#No significant orders in this phylum
```

## VERRUCOMICROBIA
```{r}
#Exclude all taxonomic metadata except for orders
verrucomicrobia_raw_abundance_withTAX_orders <- verrucomicrobia_raw_abundance_withTAX[,-c(1:3,5:7)]

#Drop all na values to only analyze OTUs assigned at the order level: 8,342total OTUs
verrucomicrobia_raw_abundance_withTAX_orders <- verrucomicrobia_raw_abundance_withTAX_orders %>% drop_na()

#Group by the same orders, and sum within each site
verrucomicrobia_orders_sums <- verrucomicrobia_raw_abundance_withTAX_orders %>%
  group_by(order) %>%
  summarize_all(funs(sum))

#Here I am inverting the dataset so that my y variables are individual orders and my x variables are sites
verrucomicrobia_orders_sum_inverted <- verrucomicrobia_orders_sums %>% column_to_rownames("order")
verrucomicrobia_orders_sum_inverted <- data.frame(t(verrucomicrobia_orders_sum_inverted))
verrucomicrobia_orders_sum_inverted <- verrucomicrobia_orders_sum_inverted %>% rownames_to_column("Site")
verrucomicrobia_orders_sum_inverted <- verrucomicrobia_orders_sum_inverted %>%
  mutate(Site = factor(Site))
verrucomicrobia_orders_sum_inverted_withTAX <- verrucomicrobia_orders_sum_inverted %>%
  left_join(site_metadata, by = "Site")

verrucomicrobia_orders_sum_inverted_withTAX <- cbind(site_metadata[-c(20:21),], verrucomicrobia_orders_sum_inverted_withTAX)
verrucomicrobia_orders_sum_inverted_withTAX <- verrucomicrobia_orders_sum_inverted_withTAX[,-c(6:7,10:12,14,17:21)]

#5 total orders used in analysis

p_vals_verrucomicrobia_orders <- data.frame("p_val" = numeric(length = 5))

    for (ii in 6:ncol(verrucomicrobia_orders_sum_inverted_withTAX)) {
  
      col = verrucomicrobia_orders_sum_inverted_withTAX[,ii] #this tells the for loop that it should iterate through columns in the data set
      ff <- as.formula(paste0(colnames(verrucomicrobia_orders_sum_inverted_withTAX)[ii]," ~ present_absent"))
      
      glmer_temp <- glm.nb(ff, data = verrucomicrobia_orders_sum_inverted_withTAX) #tells the for loop to run linear model  for each column         specified above
      p_value <- summary(glmer_temp)$coefficients[2,4] #extracts p value
      p_vals_verrucomicrobia_orders[ii-5,"p_val"] <- p_value
    }

verrucomicrobia_pval_fdr <- p.adjust(p_vals_verrucomicrobia_orders$p_val, method = "fdr")

#No significant orders in this phylum
```